# Data Sources {#data-sources}

## Introduction

Any analytic project needs access to data. The location (electronic address) or
mechanism from which data originates or can be obtained is called a *data
source*. This definition suggests data sources vary in their origins and their
storage mechanisms. First, data may originate organically, from measuring
instruments that record real-data or generated syntactically to fit a specific
purpose. Second, data may be stored inside a database, a flat-file, a website,
or an API.

The choice of what data source to use in our analytic application is taken with
regards to two major qualities: *accessibility* and *usefulness*. Given an
existing data source, accessibility is the degree of how easy it is to obtain
(access to) it. Accessibility is mainly impacted by company policy (regulation,
security protocols, etc.) and the availability and willingness of people we are
dependent on, mainly database administrators and data engineers, to respond to
our data access requests. Usefulness is the degree of having practical worth or
applicability. Usefulness is determined by circumstances combined with the goal
we are trying to achieve.

Among the two data source qualities: accessibility and usefulness, the latter
should lead our choice of a data source. The following sections describe common
goals and circumstances of analytic projects and propose appropriate data
sources.

## Data Source Usefulness

In [chapter 1](#intro), we decomposed the process of building analytic
applications into two realms: software development and data analytics. Following
that decomposition, we argue that having two data sources, one for each realm is
appropriate for many analytic projects. We begin by identifying the role of data
in each realm. Then, we infer what are the desirable or necessary attributes of
the data source in each realm. Finally, we demonstrate the proposed data sources
in our running example.

We start with data analytics as it requires the least explanations. Data
analytics is the science of analyzing data in order to make conclusions about
that information, i.e. actionable insights. Examples of data analytics
activities include data exploration and (predictive) model selection. Obviously,
when modelling the real-world, there is no substitute for real data.
Nevertheless, a representative sample of the data is sufficient and even
preferable in many analytic projects.

In our context, a representative sample is a subset of a data source that seeks
to accurately reflect the data source. Representative samples tend to contain a
subspace of the domain or a snapshot of the domain at a given time. That means
that if we turn the representative sample into actionable insights, then adding
more spatial/temporal data to the representative sample would give similar
insights.

Representative samples can be stored in flat files of modest sizes. The file
size attribute is desirable. Loading gigabytes of information every time an R
session begins is both slow and memory-intensive action. Moreover, applying
transformation on the loaded data may result in copies that would consume more
memory.

---

The following code demonstrates a syntactic data source implementation for the
running example. Notice, that the syntactic dataset captures our current
knowledge of the house prices domain. It has an identifier column (Id), salient
features (Bedrooms and Bathrooms) that must exist in the real application, and a
target variable (SalePrice). The content of the dataset holds true information
to the extent necessary for software development activities. For example, as you
would expect, SalePrice is a numeric variable with non-negative values. However,
there is no attempt (at this stage) to fake values resembling prices in the real
housing market.

```{r code = readLinesR("generate_synthetic_data.R"), echo = TRUE, eval = FALSE}
```


```{r generate_synthetic_data-exec, echo=TRUE, results='markup'}
set.seed(1356)
generate_synthetic_data(n = 100) |> head()
```


## Data Source Accessibility

## The Common Case

## Conclusions
